---
title: "Regression Models Course Project"
author: "Anjali Singh"
date: "8 October 2017"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```
## Executive Summary
In this project we will evaulate the mtcars data set and explore how miles per gallon (MPG) is affected by different variables. We will aim to answer the below two question

1) Is an automatic or manual transmission better for MPG?

2) Quantify the MPG difference between automatic and manual transmissions.

Conclusion:

Manual Transimission is much better than in Automatic Transmission in giving good mileage

1) Car with automatic transmissions gives 17.147 MPG, and with  manual transmissions car gives 24.392 MPG. 

2) In the multiple regression model, the MPG difference is 2.9358 MPG at the mean weight and qsec.

## Exploratory Analysis

```{r, results="hide"}
library(ggplot2)
data(mtcars)
summary(mtcars$mpg[mtcars$am==0])
str(mtcars)
```
Converting into factors
```{r}
mtcars$am <- as.factor(mtcars$am)
```
There are 32 observation, 10 control variable and 1 dependent varaible in this dataset, we will first try and do some exploratory data analysis to pick up any trend.
We will first check for the mean for both Auto and Manual transimission, 1 is Manual and 0 is Auto. As you see below the Mean for auto is less than that of manual. This is our first indication that Manual Transimission is better
```{r}
sapply((split(mtcars$mpg, mtcars$am)), mean)
```
Now lets try and understand this by plotting the box plot(Apendix 1)
The box plot above clearly tells us that Manual Transimission provides much better transmission than Automatic Transimission.
Lets do a t test to hypothesis to see if there is impact of Transimission type on mileage

Hypothesis:

H0:Transimission Type has no impact on Mileage

H1:Transimission Type has a impact on Mileage
```{r, results='hide'}

t.test(mtcars[mtcars$am == "0",]$mpg, mtcars[mtcars$am == "1",]$mpg)
```
As you can see, the p- Value (0.001374) is less than 0.05, therefore we reject the null hypothesis and can safely say that Transimission Type has a impact on Mileage. 

## Regression Analysis
Let us first try and do a linear model to quantify the relationship between Transimission and Mileage.

```{r,results='hide'}
lrModel <- lm(mpg ~ am, data = mtcars)
summary(lrModel)
```
THe above model only account for 36 percent of the variation in the data, which is not a good fit. This can happen if there are more variable to contributes to the variation of the model.Lets try and capture that using the Mutliple regression model

### Mulitple Regression 
We will first try and  fit the full model and try and get the predictors that are important for us to capture 
```{r,results='hide'}
initial <- lm(mpg ~ ., data = mtcars)
best <- step(initial, direction = "both")
summary(best)
```
```{r}
summary(best)
```
The three variables for the model are qsec, wt, and am, The R squared value was 84 % and Residual standard error: 2.459 on 28 degrees of freedom. P value for beta3(am) is 0.047 and is small enough for us to reject the null hypothesis and safely say that Transmission does impact mileage.

### Residuals and Diagnostics(Appendix 2)

The Normal Q-Q plot indicates the normal distribution of the date while residual vs fitted shows the data us random and and is independent.
Scale location plot indicates  spread of points across predicted values range and here the residuals have uniform variance across the range.There is no Homoscedasticity
The residuals for the Chrysler Imperial, Fiat 128, and Toyota Corolla  are outliers and exert some influence on the curve



### Apendices
#### Apendix 1
```{r}
#Visualization ~ Automatic vs Manual Transmission:
library(ggplot2)

g <- ggplot(aes(x=am, y=mpg), data=mtcars) + geom_boxplot(aes(fill=am))
g <- g + labs(title = "Automatic(0)vs Manual(1) Trans Boxplot")
g <- g + xlab("Transmission Type")
g
```

#### Apendix 2
```{r}
par(mfrow=c(2,2))
plot(best)
```

